Reduced growth due to belowground sink limitation is not fully explained by reduced photosynthesis

### 2016-11-04 manuscript/manuscript_pve.Rmd
tiny fixes front page
last edits

\
Courtney E. Campany^1^, Belinda Medlyn^1^ and Remko A. Duursma^1^

@@

Runnging Head: Sink limitation and reduced seedling growth

@@

Sink limitation is known to reduce plant growth, but it is not known how plant carbon (C) balance is affected, limiting our ability to predict growth under sink-limited conditions. We manipulated soil volume to impose sink limitation of growth in *Eucalyptus tereticornis* seedlings. Seedlings were grown in the field in containers of different sizes and planted flush to the soil alongside freely-rooted (Free) seedlings. Container volume negatively affected aboveground growth throughout the experiment, and light saturated rates of leaf photosynthesis were consistently lower in seedlings in containers (-26%) compared to Free seedlings. Photosynthetic capacity was also significantly reduced in containerized seedlings, and was related to both reduced leaf nitrogen content and starch accumulation.  After 120 days, harvested biomass of Free seedlings was on average 84% higher than seedlings in containers, but biomass distribution in leaves, stems and roots was not different.
 However, the reduction in net leaf photosynthesis over the growth period was insufficient to explain the reduction in growth, so that we also observed an apparent reduction in whole-plant C use efficiency (CUE) with container size. Our results show that sink limitation affects plant growth through feedbacks to both photosynthesis and CUE. Mass balance approaches to predicting plant growth under sink-limited conditions need to incorporate both of these feedbacks.

@@

\
Total seedling biomass at harvest was significantly different across container volumes (*P* < 0.001) and between container treatments and Free seedlings (*P* < 0.001, Table `r pvetab1`). On average, harvested biomass of Free seedlings was 84% higher than that of seedlings in containers. Plant biomass was positively correlated with total leaf area across all treatments (*R*^2^ = 0.97, *P* < 0.001). Differences in biomass partitioning to leaves, stems and roots were not different across soil volumes after variation in seedling biomass across treatments was factored in the analysis (Figure `r allocfig`a,b). The final harvest root:shoot biomass ratio and  was conserved across all treatments, with a slightly higher shoot than root mass ($\overline{x}$ = 0.904, 95% CI = [0.846,1.119]). The ratio of leaf to fine root mass was also not different between treatments (Figure `r allocfig`c).

>>>>

\
Courtney E. Campany^1^, Belinda E. Medlyn^1^ and Remko A. Duursma^1^

@@

Running Head: Sink limitation and reduced seedling growth

@@

Sink limitation is known to reduce plant growth, but it is not known how plant carbon (C) balance is affected, limiting our ability to predict growth under sink-limited conditions. We manipulated soil volume to impose sink limitation of growth in *Eucalyptus tereticornis* seedlings. Seedlings were grown in the field in containers of different sizes and planted flush to the soil alongside freely-rooted (Free) seedlings. Container volume negatively affected aboveground growth throughout the experiment, and light saturated rates of leaf photosynthesis were consistently lower in seedlings in containers (-26 %) compared to Free seedlings. Photosynthetic capacity was also significantly reduced in containerized seedlings, and was related to both reduced leaf nitrogen content and starch accumulation.  After 120 days, harvested biomass of Free seedlings was on average 84 % higher than seedlings in containers, but biomass distribution in leaves, stems and roots was not different. However, the reduction in net leaf photosynthesis over the growth period was insufficient to explain the reduction in growth, so that we also observed an apparent reduction in whole-plant C use efficiency (CUE) with container size. Our results show that sink limitation affects plant growth through feedbacks to both photosynthesis and CUE. Mass balance approaches to predicting plant growth under sink-limited conditions need to incorporate both of these feedbacks.

@@

\
Total seedling biomass at harvest was significantly different across container volumes (*P* < 0.001) and between container treatments and Free seedlings (*P* < 0.001, Table `r pvetab1`). On average, harvested biomass of Free seedlings was 84 % higher than that of seedlings in containers. Plant biomass was positively correlated with total leaf area across all treatments (*R*^2^ = 0.97, *P* < 0.001). Differences in biomass partitioning to leaves, stems and roots were not different across soil volumes after variation in seedling biomass across treatments was factored in the analysis (Figure `r allocfig`a,b). The final harvest root:shoot biomass ratio and  was conserved across all treatments, with a slightly higher shoot than root mass ($\overline{x}$ = 0.904, 95 % CI = [0.846,1.119]). The ratio of leaf to fine root mass was also not different between treatments (Figure `r allocfig`c).
